---
title: "Replication of analyses reported in Montero-Melis & Jaeger (2019) in *BLC*"
author: '[Guillermo Montero-Melis](https://www.mpi.nl/people/montero-melis-guillermo)'
date: '`r as.character(format(Sys.Date(), format="%d/%m/%Y"))`'
output:
  html_document:
    depth: 2
    number_sections: yes
    theme: default
    toc: yes
---



Set up working space
====================

Load libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 4, fig.width = 5)
```

```{r, include = TRUE, echo = TRUE, warning = FALSE, message = FALSE}
library(dplyr)
packageVersion('dplyr')
# library(tidyr)
# packageVersion('tidyr')  # useful for gather() to convert from wide to long
# library(mgcv)  # GAMs and GAMMs (Wood 2006)
# packageVersion('mgcv')
# library(itsadug)
# packageVersion('itsadug')  # Visualizing GAMMs
# library(lme4)
# packageVersion('lme4')
# library(ggplot2)
# packageVersion('ggplot2')
# library(boot)  # for inv.logit()
# packageVersion('boot')
# library(knitr)  # for kable()
# packageVersion('knitr')
# library(lazyeval)  # lazy evaluation used in bysubj() function [summarise_]
# packageVersion('lazyeval')
# library(effects)
# library(xtable)
```


Load data
=========

```{r}
# Make sure you are in the folder where all the scripts and data are located
getwd()
```


```{r}
# load data
d <- read.csv('data.csv', fileEncoding = 'UTF-8', stringsAsFactors = TRUE)
# Make sure Subject is a factor
d$Subject <- factor(d$Subject)
# in Group, let the native speakers (NS) be the reference group
d$Group <- factor(d$Group, levels = c('NS', 'L2'))
head(d, 3)
# load participant information 
# NB: Accidentally, no audio data was recorded for Subject 14 (L2), so this
# participant is omitted from all participant descriptors as well.
ppts <- read.csv("participant-info.csv", fileEncoding = "UTF-8", 
                 stringsAsFactors = TRUE)
head(ppts, 3)
tail(ppts, 3)  # L2 proficiency-related variables only available for L2ers
```


Participant descriptors
=======================

## As reported in the main text (Method section)

Number of participants by group:

```{r}
# Number of participants by group that go into the analysis (note data for
# one L2 speaker was missing because of recording failure)
table(ppts$Group)
```


Participant age:

```{r}
# age
ppts %>%
  group_by(Group) %>%
  summarise(Mage = mean(Age, na.rm = T), SDage = sd(Age, na.rm = T))
```


Learners' age of onset for learning Spanish 

```{r}
# Mean and SD
ppts %>%
  filter(Group == "L2") %>%
  summarise(M  = round(mean(L2_AoO, na.rm = TRUE), 2), 
            SD = round(sd(L2_AoO, na.rm = TRUE), 2))
# range
range(ppts$L2_AoO, na.rm = TRUE)
```



Scores on offline cloze test ('proficiency score'):

```{r}
# Cloze scores
round(mean(ppts$ClozeScore, na.rm=T), 1)
round(sd(ppts$ClozeScore, na.rm=T), 1)
# By condition
ppts %>% 
  filter(Group == "L2") %>% 
  group_by(Condition) %>%
  summarise(M_profic = mean(ClozeScore), 
            SD = sd(ClozeScore))
```


Difference in L2 proficiency across conditions?

```{r}
ppts %>% 
  filter(Group == "L2") %>%
  aov(ClozeScore ~ Condition, data = .) %>%
  summary
```


Computerized vocabulary task (deliberately easy)

```{r}
# Scores on vocabulary task
round(mean(ppts$VocabTaskAcc, na.rm=T), 3)
round(sd(ppts$VocabTaskAcc, na.rm=T), 4)
# By condition
ppts %>% 
  filter(Group == "L2") %>% 
  group_by(Condition) %>%
  summarise(M_profic = mean(VocabTaskAcc), 
            SD = sd(VocabTaskAcc))
```



## Participant information reported in the Supplementary Online Information (S2)


```{r}
with(ppts, table(Group, Gender))
with(ppts, table(Group, Condition))
# age
ppts %>%
  group_by(Group) %>%
  summarise(Mage = mean(Age, na.rm = T), SDage = sd(Age, na.rm = T))
# Cloze scores
mean(ppts$ClozeScore, na.rm=T)
sd(ppts$ClozeScore, na.rm=T)
# By condition
ppts %>% filter(Group == "L2") %>% group_by(Condition) %>%
  summarise(M_profic = mean(ClozeScore), SD = sd(ClozeScore))
```

Session info
============

```{r}
sessionInfo()
```


