---
title: "Analysis L2 lexicalisation"
author: '[Guillermo Montero-Melis](http://www.biling.su.se/montero_melis_guillermo)'
date: '`r as.character(format(Sys.Date(), format="%d/%m/%Y"))`'
output:
  pdf_document:
    number_sections: yes
    toc: yes
  html_document:
    depth: 2
    number_sections: yes
    theme: default
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.height = 4, fig.width = 5)
```


# Introduction

During production, speakers adapt to the patterns experienced in recent linguistic exposure.
Here we probe the difference between native speakers (NS) of Spanish and Swedish second language speakers (L2S) of Spanish regarding whether and how they adaptat their lexicalisation preferences (use of path- vs. manner-verbs) in the course of a production elicitation task.
All participants had to describe caused motion events (e.g., 'He pushed a box up a sand dune').
Participants in the Path-exposed condition read out loud sentences containing a path verb that could be used to describe the target event
(e.g., *El se침or sube unos escalones con una televisi칩n* 'The man ascends some steps with a television'), 
whereas participants in the Manner-exposed condition read sentences with manner verbs
(e.g., *El se침or empuja una televisi칩n por unos escalones* 'The man pushes a television along some steps').
Their respective lexicalisation preferences were compared to those of control groups (one for NS and one for L2S) who did not receive any exposure while describing the events.


## Research questions

1. Is adaptation overall stronger in L2S than NS?
2. Is adaptation in L2S mediated by their first language (L1) biases?
3. Does the time course of adaptation differ between the two groups?
4. Do L2S' patterns of adaptation come to resemble more those of NS as their proficiency increases?


## Predictions

1. Because L2S have overall less experience in their L2 than NS, their production patterns should be more malleable and thus they should adapt more as they are exposed to sentences in the target language.
2. Because manner verbs constitutes the preferred lexicalisation pattern in Swedish (i.e., L2 speakers' L1), L2 speakers should adapt less in the manner-exposure condition relative to NS, while they should adapt more in the path-exposure condition relative to NS.
3. We expected adaptation to be relatively fast in the high-surprise (=unexpected) conditions, i.e. in the path-exposure condition for L2 speakers and in the manner-exposure condition for NS; whereas we expected slower adaptation in the low-surprise (=expected) conditions, i.e. manner-exposure for L2 speakers and path-exposure for NS.
4. As they become more proficient in the target language, L2 speakers should be less influenced by their L1-based lexicalisation preferences and have developed a more stable preference for the Spanish pattern. Thus, we expected that with increasing proficiency L2 speakers would adapt less to path-exposure and adapt more to manner-exposure.


# Set up working space

## Libraries

```{r, include = TRUE, echo = TRUE, warning = FALSE, message = FALSE}
library(dplyr)
packageVersion('dplyr')
library(mgcv)  # GAMs and GAMMs (Wood 2006)
packageVersion('mgcv')  # GAMs and GAMMs (Wood 2006)
library(itsadug)
packageVersion('itsadug')
library(lme4)
packageVersion('lme4')
library(ggplot2)
packageVersion('ggplot2')
library(boot)  # for inv.logit()
packageVersion('boot')  # for inv.logit()
library(knitr)  # for kable()
packageVersion('knitr')  # for kable()
library(lazyeval)  # lazy evaluation used in bysubj() function [summarise_]
packageVersion('lazyeval')
```


## Data loading and processing

### Load data

```{r}
##  Load data
# The data is created in the script 'processing/compute_dependent_measures.R'
# There is the normal and the liberally coded version, see script for difference

# load
d <- read.csv('../data/data_DVs.csv', fileEncoding = 'UTF-8', stringsAsFactors = TRUE)
# simplify somewhat
d <- d %>%
  select(Subject:VideoName, P_V, M_V) %>%
  rename(Trial = VideoTrial)
# head(d)
# str(d)

# participant data
ppts <- read.csv("../data/participants.csv", fileEncoding = "UTF-8", stringsAsFactors = TRUE)
# head(ppts)
# No audio data recorded for Subject 14 (L2) due to experimental error; exclude
ppts <- ppts[ppts$Subject != 14, ]

# add speakers' clozescore to d:
d <- left_join(d, ppts %>% select(Subject, ClozeScore))

# Subject needs to be a factor
d$Subject <- factor(d$Subject)
# in Group, let the native speakers (NS) be the reference group
d$Group <- factor(d$Group, levels = c('NS', 'L2'))

head(d)
str(d)
```


### Remove outliers?

Remove any L2 speakers based on cloze score because too low?

```{r}
# Outliers? ---------------------------------------------------------------

# L2 speakers Cloze scores
cloze <- ppts %>%
  filter(Group == "L2") %>%
  select(Subject, Condition, ClozeScore)
# z-scores
cloze$zClozeScore <- scale(cloze$ClozeScore)
# Subjects ordered by clozescore
head(cloze[order(cloze$zClozeScore), ])
tail(cloze[order(cloze$zClozeScore), ])

# Consider ClozeScore < 5 as outlier
threshold5 <- cloze[cloze$ClozeScore < 5, "Subject"]
# # Consider ClozeScore < 10 as outlier
threshold10 <- cloze[cloze$ClozeScore < 10, "Subject"]

## Exclude participants? (Uncomment to exclude)
# d <- d[!d$Subject %in% threshold5, ]
# d <- d[!d$Subject %in% threshold10, ]
```

### Data for GAMs comparing NS and L2 speakers

Subset data, define factors, set coding scheme, etc:

```{r}
## Single factor to model Group * Condition, using dummy coding:
# (The way to analyze crossed factors seems to be to combine their levels into 
# a single factor, see Wood's comment here
# http://grokbase.com/t/r/r-help/113qaadxt4/r-how-to-add-in-interaction-terms-in-gamm)
d$GroupCondit <- with(d, interaction(Group, Condition))

## Subset data:
# Data for analysis of Path-priming vs control
dp <- d %>% filter(Condition != 'Manner')
dp$Condition <- factor(dp$Condition) # drop unused levels
# Data for analysis of Manner-priming vs control
dm <- d %>% filter(Condition != 'Path')
dm$Condition <- factor(dm$Condition) # drop unused levels

# Order levels correctly in each data frame
dp$GroupCondit <- factor(dp$GroupCondit, levels = c("NS.Control", "NS.Path",
                                                    "L2.Control", "L2.Path"))
dm$GroupCondit <- factor(dm$GroupCondit, levels = c("NS.Control", "NS.Manner",
                                                    "L2.Control", "L2.Manner"))
# note that NS control becomes the reference group in both data frames
contrasts(dp$GroupCondit)
contrasts(dm$GroupCondit)
```


### Data for GAMs testing effect of proficiency in L2 speakers

Subset data, define factors, set coding scheme, etc:

```{r}
# subset L2 data only from Path/Control and Manner/Control conditions respectively
dp_l2 <- dp %>% filter(Group == "L2")
dm_l2 <- dm %>% filter(Group == "L2")
```


### Data for logit mixed models

Subset data, define factors, set coding scheme, etc:

```{r, echo = T}
## Data sets for mixed model fitting
dp_lme4 <- dp
dm_lme4 <- dm
# We use a coding that allows us to discover any effects of exposure across groups
contrasts(dp_lme4$Group) <- - contr.sum(2) / 2
colnames(contrasts(dp_lme4$Group)) <- "L2_vs_NS"
contrasts(dp_lme4$Group)
contrasts(dm_lme4$Group) <- - contr.sum(2) / 2
colnames(contrasts(dm_lme4$Group)) <- "L2_vs_NS"
contrasts(dm_lme4$Group)
# for condition use dummy coding with Control as reference
contrasts(dp_lme4$Condition)
contrasts(dm_lme4$Condition)
```



## Convenience functions etc.

[NB: might be hidden if `echo = FALSE`.]

```{r}
## Specify some global parameters

# adjust figure heght/width when not going with default (espec. for 2x2 plots)
myfighe_NS_L2 <- 6
myfighe_L2_prof <- 6
myfigwi <- 7
```


```{r}
## Convenience functions

# source functions to compute collinearity diagnostics
source("functions/kappa_mer_fnc.R")
source("functions/vif_mer_fnc.R")

# print deviance explained as percentage
dev_expl <- function(fm) {
  devi <- summary(fm)$dev.expl
  paste0(round(100 * devi, 1), '% dev. explained')
}

# create a table of the summary of fixed effects of a mixed model
glmm_tb <- function(fm) {
  m <- round(summary(fm)$coefficients, 3)
  tb <- as.data.frame(m)
  names(tb) <- c("Estimate", "SE", "z-value", "p-value")
  kable(tb)
}

# Function used to load models if they have already been saved,
# rather than fitting them anew each time the script is called
load_or_fit <- function(fm_name, fm_code, forcefit = FALSE) {
  rel_path <- './gamms/'
  # the string may contain '\n'
  fm_code <- gsub('\n', '', fm_code)
  
  # check first if there is a saved rda object with that name; if so, load it
  model_exists <- paste0(fm_name, '.rda') %in% list.files(path = rel_path)
  # if forcefit = T it will always fit the model
  if((!forcefit) & model_exists) {
    load(paste0(rel_path, fm_name, '.rda'), .GlobalEnv)
    
  } else {  # fit it and save it to right folder
    # record the time it takes to fit
    fm_time <- paste0(fm_name, ".t")
    ptm <- proc.time()
    assign(fm_name, eval(parse(text = fm_code)), envir = .GlobalEnv)
    assign(fm_time, proc.time() - ptm, envir = .GlobalEnv)
    # both the model and the time it took can be saved as same file
    save(list = c(fm_name, fm_time), file = paste0(rel_path, fm_name, '.rda'))
    # print a brief model summary when it has been fitted and time to fit
    fm <- get(fm_name)
    fm_t <- get(fm_time)
    print(fm)
    print(fm_t)
  }
}

# function to plot the differences between native and L2 speakers
plot_NS_L2 <- function(fm, primed_cond = NULL, ylim1 = c(-4, 6), ylim2 = c(-.5, 6), ...) {
  layout(matrix(1:4, ncol = 2, byrow=TRUE), heights = c(1.5, 1))
  par(mai = c(.7, .8, .5, 0.2))
  mycex <- 0.9
  NS_primed <- paste0('NS.', primed_cond)
  L2_primed <- paste0('L2.', primed_cond)
  # plot NS control
  plot_smooth(fm, view = 'Trial', cond = list(GroupCondit = 'NS.Control'),
              col = 'blue', lty = 2, rug = FALSE, ylim = ylim1, rm.ranef=TRUE,
              main = 'Native speakers', ylab = paste0('Log-likelihood\nof ', primed_cond, " verb"),
              hide.label = TRUE)
  # plot NS primed
  plot_smooth(fm, view = 'Trial', cond = list(GroupCondit = NS_primed),
              col = 'red', rug = FALSE, rm.ranef = TRUE, add = TRUE,
              hide.label = TRUE)
  # add legend
  legend(x = 1, y = ylim1[2] + 0.75, legend = c(paste0(primed_cond, '-primed'), 'Control'),
         col = c('red', 'blue'), lty = 1:2, cex = mycex, box.lty = 0)
  
  # L2 speakers Control
  plot_smooth(fm, view = 'Trial', cond = list(GroupCondit = 'L2.Control'),
              col = 'blue', lty = 2, rug = FALSE, ylim = ylim1, rm.ranef = TRUE,
              main = 'L2 speakers', ylab = paste0('Log-likelihood\nof ', primed_cond, " verb"),
              hide.label = TRUE)
  # L2 speakers primed
  plot_smooth(fm, view = 'Trial', cond = list(GroupCondit = L2_primed),
              col = 'red', rug = FALSE, rm.ranef = TRUE, add = TRUE,
              hide.label = TRUE)
  # add legend
  legend(x = 1, y = ylim1[2] + 0.75, legend = c(paste0(primed_cond, '-primed'), 'Control'),
         col = c('red', 'blue'), lty = 1:2, cex = mycex, box.lty = 0)
  
  # Now plot the estimated differences with itsadug::plot_diff()
  plot_diff(fm, view = 'Trial', comp = list(GroupCondit = c(NS_primed, 'NS.Control')),
            ylim = ylim2, rm.ranef=TRUE, 
            main = paste0('NS: ', primed_cond, '-primed',  ' - Control'),
            ylab = paste('Diff. in log-likelihood\nof ', primed_cond, 'verb'), hide.label = TRUE,
            ...)  # a hack I need to set the mark.diff argument as F, bc of some bug?
  plot_diff(fm, view = 'Trial', comp = list(GroupCondit = c(L2_primed, 'L2.Control')),
            ylim = ylim2, rm.ranef=TRUE, 
            main = paste0('L2: ', primed_cond, '-primed',  ' - Control'),
            ylab = paste('Diff. in log-likelihood\nof ', primed_cond, 'verb'), hide.label = TRUE)
}


# function to plot the effects by L2 speakers' proficiency
plot_L2_profic <- function(fm, primed_cond = NULL, ylim1 = c(-6, 8),
                           cloze_range = c(6, 36)) {
  cloze_scores <- seq(cloze_range[1], cloze_range[2], length.out = 6)
  # we want to have one common legend for all plots
  layout(rbind(1, matrix(2:7, ncol=3, byrow=TRUE)), heights = c(1, 6, 6))
  # add common legend
  par(mai=c(0,0,0,0))
  plot.new()
  legend(x = "center", ncol = 2, legend = c(paste0(primed_cond, '-primed'), 'Control'),
         col = c('red', 'blue'), lty = 1:2, box.lty = 0, cex = 1.5)
  # Now make plot for the model estimates at the different cloze scores in cloze_scores
  # par(mai=rep(0.2, 4))
  par(mai = c(.7, .8, .5, 0))
  for(cloze in cloze_scores) {
    plot_smooth(fm, view = 'Trial', cond = list(Condition = 'Control', ClozeScore = cloze),
                col = 'blue', lty = 2, rug = FALSE, ylim = ylim1, rm.ranef = TRUE,
                hide.label = TRUE, ylab = paste0('Log-likelihood\nof ', primed_cond, " verb"),
                main = paste('Cloze score =', cloze))
    plot_smooth(fm, view = 'Trial', cond = list(Condition = primed_cond, ClozeScore = cloze),
                col = 'red', rug = FALSE, rm.ranef = TRUE, hide.label = TRUE, add = TRUE)
    }
}
```


```{r}
## Source a function that plots model-estimates in logit space with speaker estimates!
source("functions/plot_glmm_fnc.R")
```



# Analyses and results

## Aggregated effects of adaptation (trial not considered)


### Path verbs in path-exposed vs control


The log-likelihood of using a path verb (`P_V`) is modelled as a function of:

- Group (L2 vs NS)
- Condition (Path-exposed vs Control)
- Including random intercepts by speaker and item (random slopes by item were not justified by the data)


Model fit and summary of fixed effects:

```{r, echo = TRUE}
glmm_pve <- glmer(P_V ~ Group * Condition + (1 | Subject) + (1 | VideoName),
                data = dp_lme4, family = "binomial", 
                control = glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
glmm_tb(glmm_pve)
```


```{r}
plot_glmm(glmm_pve, d = dp_lme4, exposed = "Path", ylims = c(-5.5, 5))
```


**Model interpretation against predictions**

- L2 speakers used path verbs to the same extent as NS in the control condition: both groups used path verbs slightly more often than half of the time.
- Furthermore, there was a main effect of path-exposure, leading to more path verbs when exposed to path sentences than in the control condition.
- However, there was no interaction of group and path-exposure, suggesting that both groups adapted to similar extents their production preferences.


More detail about the model:

```{r}
summary(glmm_pve)
kappa.mer(glmm_pve)
vif.mer(glmm_pve)
```



### Manner verbs in Manner-exposed vs control


The log-likelihood of using a Manner verb (`M_V`) is modelled as a function of:

- Group (L2 vs NS)
- Condition (Manner-exposed vs Control)
- Including random intercepts by speaker and item (random slopes by item were not justified by the data)


Model fit and summary of fixed effects:

```{r, echo = TRUE}
glmm_mve <- glmer(M_V ~ Group * Condition + (1 | Subject) + (1 | VideoName),
                data = dm_lme4, family = "binomial", 
                control = glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
glmm_tb(glmm_mve)
```

```{r}
# plot model predictions
plot_glmm(glmm_mve, d = dm_lme4, exposed = "Manner", ylims = c(-5.5, 5))
```


**Model interpretation against predictions**

- In the control condition, L2 speakers used manner verbs less often than NS. 
NS had an average probability of
`r round(inv.logit(summary(glmm_mve)$coefficients[1,1]), 2)`
of producing a manner verb,
while L2 speakers had an average probability of `r round(inv.logit(sum(summary(glmm_mve)$coefficients[1:2,1])), 2)`.
- There was a main effect of manner-exposure, leading to more naber verbs when exposed to manner sentences compared to the control condition.
- Again, there was no interaction of group and manner-exposure, suggesting that both groups adapted to similar extents their production preferences.




More detail about the model:

```{r}
summary(glmm_mve)
kappa.mer(glmm_mve)
vif.mer(glmm_mve)
```



### Summary of results





## Differences in adaptation over time (trial-by-trial analysis using GAMs)


### Path verbs in path-exposed vs control -- trial-by-trial

Previous models (not shown here) showed that random *intercepts* by items made sense, but not random slopes.
We add the former but not the latter.

Thus, the model is specified as:

`PathVerb ~ GroupCondit + s(Trial, by = GroupCondit) + s(Trial, Subject, bs = 'fs') + s(VideoName, bs = 're')`

Predictors are:

- `GroupCondit`: Group-Condition interaction as a fixed effect
- `s(Trial, by = GroupCondit)`: A smooth function of `Trial` allowing the function to differ for each level of Group-Condition (thin plate regression splines, the default)
- `s(Trial, Subject, bs = 'fs')`: Factor smooths for `Subject` to capture non-linear random effects of speakers
- `s(VideoName, bs = 're')`: Random intercepts by items (i.e., `VideoName`)


```{r}
# the expression that is passed to load_or_fit()
fm_p3.expr <- "bam(P_V ~ GroupCondit + s(Trial, by = GroupCondit) +
                    s(Trial, Subject, bs = 'fs') + s(VideoName, bs = 're'),
                  data = dp,
                  family = 'binomial')"
# load model or fit
load_or_fit("fm_p3", fm_p3.expr)
```

Summary of the model:

```{r}
summary(fm_p3)
```

```{r}
# kable(gamtabs(fm_p3))
```



Significance of the different terms in the model:

```{r}
anova(fm_p3)
```


Plot model estimates by condition, random smooth adjustments by speakers and QQ-plot of random by-item intercepts:

```{r, results = 'hide'}
par(mfrow = c(1, 1))
## plot
plot_smooth(fm_p3, view = 'Trial', plot_all = 'GroupCondit', rm.ranef=TRUE)
```


```{r}
# show by-speaker random smooths and by-item random intercepts
par(mfrow = c(1, 2))
plot(fm_p3, select = 5)
plot(fm_p3, select = 6)
```



Plot differences between conditions (model estimates):

```{r, fig.height = myfighe_NS_L2, fig.width = myfigwi, echo = FALSE, results = 'hide'}
plot_NS_L2(fm_p3, primed_cond = 'Path', mark.diff = FALSE)  # mark.diff is a hack to
# avoid an error that will stop the whole thing when I compile it on my work pc
# (it works fine on my own pc with linux)
```


### Manner verbs in Manner-exposed vs control -- trial-by-trial

Previous models (not shown here) showed that random *intercepts* by items made sense, but not random slopes.
We add the former but not the latter.

Thus, the model is specified as:

`MannerVerb ~ GroupCondit + s(Trial, by = GroupCondit) + s(Trial, Subject, bs = 'fs') + s(VideoName, bs = 're')`

Predictors are:

- `GroupCondit`: Group-Condition interaction as a fixed effect
- `s(Trial, by = GroupCondit)`: A smooth function of `Trial` allowing the function to differ for each level of Group-Condition (thin plate regression splines, the default)
- `s(Trial, Subject, bs = 'fs')`: Factor smooths for `Subject` to capture non-linear random effects of speakers
- `s(VideoName, bs = 're')`: Random intercepts by items (i.e., `VideoName`)


```{r}
# the expression that is passed to load_or_fit()
fm_m3.expr <- "bam(P_V ~ GroupCondit + s(Trial, by = GroupCondit) +
                    s(Trial, Subject, bs = 'fs') + s(VideoName, bs = 're'),
                  data = dm,
                  family = 'binomial')"
# load model or fit
load_or_fit("fm_m3", fm_m3.expr)
```

Summary of the model:

```{r}
summary(fm_m3)
```


Significance of the different terms in the model:

```{r}
anova(fm_m3)
```


Plot model estimates by condition, random smooth adjustments by speakers and QQ-plot of random by-item intercepts:

```{r, results = 'hide'}
## plot
plot_smooth(fm_m3, view = 'Trial', plot_all = 'GroupCondit', rm.ranef=TRUE)
```


```{r}
# show by-speaker random smooths and by-item random intercepts
par(mfrow = c(1, 2))
plot(fm_m3, select = 5)
plot(fm_m3, select = 6)
```



Plot differences between conditions (model estimates):

```{r, fig.height = myfighe_NS_L2, fig.width = myfigwi, echo = FALSE, results = 'hide'}
plot_NS_L2(fm_m3, primed_cond = 'Manner')
```



## Changes in adaptation with increasing L2 proficiency


### Adaptation to path verbs


```{r}
fm_p_prof9.expr <- "bam(P_V ~ Condition +
                    te(Trial, ClozeScore, by = Condition) +
                    s(Trial, Subject, bs = 'fs') +
                    s(VideoName, bs = 're'),
                  data = dp_l2,
                  family = 'binomial')"
# load model or fit
load_or_fit("fm_p_prof9", fm_p_prof9.expr)
summary(fm_p_prof9)
```


```{r}
# show by-speaker random smooths and by-item random intercepts
par(mfrow = c(1, 2))
plot(fm_p_prof9, select = 3)
plot(fm_p_prof9, select = 4)
```



```{r, fig.height = myfighe_L2_prof, fig.width = myfigwi, results='hide'}
plot_L2_profic(fm_p_prof9, primed_cond = 'Path')
```


```{r, results='hide'}
# difference plot in 3D
plot_diff2(fm_p_prof9, view = c('Trial', 'ClozeScore'), 
           comp = list(Condition = c('Path', 'Control')),
           rm.ranef = TRUE)
```



### Adaptation to manner verbs


If we consider a model analogous to the one we fitted for path, we see the trends we expected.
That is, *numerically* the trends follow the predictions.
However, none of the terms related to `ClozeScore` are actually significant in the model -- not as fixed effect, either simple or in interaction; not as smooth function, whether it is allowed to vary by condition or not, whether it interacts with trial or not.
In brief, the data don't support an effect of proficiency (as measured by `ClozeScore`) on adaptation to manner.


```{r}
fm_m_prof9.expr <- "bam(M_V ~ Condition +
                    te(Trial, ClozeScore, by = Condition) +
                    s(Trial, Subject, bs = 'fs') +
                    s(VideoName, bs = 're'),
                  data = dm_l2,
                  family = 'binomial')"
# load model or fit
load_or_fit("fm_m_prof9", fm_m_prof9.expr)
summary(fm_m_prof9)
```

```{r}
# show by-speaker random smooths and by-item random intercepts
par(mfrow = c(1, 2))
plot(fm_m_prof9, select = 3)
plot(fm_m_prof9, select = 4)
```


```{r, fig.height = myfighe_L2_prof, fig.width = myfigwi, results='hide'}
plot_L2_profic(fm_m_prof9, primed_cond = 'Manner')
```

```{r, results='hide'}
# difference plot in 3D
plot_diff2(fm_m_prof9, view = c('Trial', 'ClozeScore'), 
           comp = list(Condition = c('Manner', 'Control')), rm.ranef = TRUE)
```

