---
title: 'What is the best way to bin L2 participants by proficiency?'
author: '[Guillermo Montero-Melis](http://www.biling.su.se/montero_melis_guillermo)'
date: '`r as.character(format(Sys.Date(), format="%d/%m/%Y"))`'
output:
  html_document:
    depth: 2
    number_sections: yes
    theme: default
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document:
    toc: yes
---

```{r, message = FALSE}
library(dplyr)
library(ggplot2)
```


Aim
===

What is the best way to bin L2 participants by proficiency?

 We will loop through different values and try to maximize:
 
- The number of participants in low vs high proficiency groups and in each condition
- The mean difference in Cloze score between the groups

We'll do this by visual inspection.


Load data
=========

```{r}
d <- read.csv("../data/data_verbs_multinomial.csv")
# head(d)
# unique L2 participants and their Cloze scores:
l2 <- unique(d[, c("Subject", "Group", "Condition", "ClozeScore")])
l2 <- l2[l2$Group == "L2", ]
head(l2)
```


Distribution of proficiency scores
===================================


```{r}
ggplot(l2, aes(x = ClozeScore, fill = Condition)) +
  geom_density(alpha = .5) 
```

I had previously chosen the 4th and 6th deciles as cutoff points for low and
high proficiency speakers, respectively:

```{r}
cutoff_prof <- quantile(l2$ClozeScore, c(.4, .6), na.rm = TRUE)
cutoff_prof
```


```{r}
ggplot(l2, aes(x = ClozeScore, fill = Condition)) +
  geom_density(alpha = .5) +
  geom_vline(xintercept = cutoff_prof) +
  annotate("text", x = 10, y = .055, label = "low\nproficiency") +
  annotate("text", x = 33, y = .055, label = "high\nproficiency")
```



Number of observations depending on threshold
=============================================

Calculate the precise number of observations we have in each condition
depending on what we choose as the threshold for:

- `low_profic`: meaning that a participant with ClozeScore *equal or below* that
value will be assigned to the **low proficiency** group
- `high_profic`: meaning that a participant with ClozeScore *equal or above* that
value will be assigned to the **high proficiency** group


```{r}
compute_n <- function(df = l2) {
  my_n <- data.frame()
  min_profic <- min(df$ClozeScore)
  max_profic <- max(df$ClozeScore)
  for (thresh in seq(min_profic, max_profic)) {
    curr_low <- df %>% 
      filter(ClozeScore <= thresh) %>%
      group_by(Condition) %>%
      summarise(N = n())
    curr_low$Threshold_type <- "low_profic"
    curr_high <- df %>% 
      filter(ClozeScore >= thresh) %>%
      group_by(Condition) %>%
      summarise(N = n())
    curr_high$Threshold_type <- "high_profic"
    curr <- rbind(curr_low, curr_high)
    curr$Threshold <- thresh
    my_n <- rbind(my_n, curr)
  }
  my_n
}
my_n <- compute_n()
head(my_n)
```


Plot the results

```{r, fig.width = 10}
p <- ggplot(my_n, aes(x = Threshold, y = N, colour = Condition,
                      linetype = Threshold_type, shape = Threshold_type)) +
  geom_point() +
  geom_line()
p
```


Mean difference depending on threshold
======================================

```{r}
min_low_profic_thresh <- 21  # minimum proficiency threshold for low profic speakers
max_high_profic_thresh <- 28  # maximum proficiency threshold for high profic speakers
```


The previous plot already sets a limit on the range of thresholds we can consider.
Assuming we don't want there to be less than 6 participans per condition and
proficiency group...

- We need a `low_profic` theshold of at least `r min_low_profic_thresh`
- We need a `high_profic` theshold of at most `r max_high_profic_thresh`

Compute the mean difference in proficiency score as a function of

- `low_profic`: meaning that a participant with ClozeScore *equal or below* that
value will be assigned to the **low proficiency** group
- `high_profic`: meaning that a participant with ClozeScore *equal or above* that
value will be assigned to the **high proficiency** group

Note that:

- `low_profic` $<$ `high_profic`: the threshold for low proficiency has to be
strictly smaller than the threshold for high proficiency
- However, there can be a gap, i.e. we could set the lower threshold at 22 and
the upper at 25, thus losing all participants with proficiency in the range [23, 24]

```{r}
compute_diff <- function(df = l2,
                         # minimum proficiency threshold for low profic speakers
                         min_low_profic = min_low_profic_thresh,
                         # maximum proficiency threshold for high profic speakers
                         max_high_profic = max_high_profic_thresh) {
  mean_diff <- data.frame()  # initialize empty data frame
  for (lower_thresh in min_low_profic : (max_high_profic - 1)){
    M_low <- mean(df[df$ClozeScore <= lower_thresh, "ClozeScore"])
    for (upper_thresh in (lower_thresh + 1) : max_high_profic) {  # upper threshold strictly greater than lower
      curr <- df %>% 
        filter(ClozeScore >= upper_thresh) %>%
        summarise(M_high = mean(ClozeScore))
      curr$Threshold_high <- upper_thresh
      curr$Threshold_low <- lower_thresh
      curr$M_low <- M_low
      mean_diff <- rbind(mean_diff, curr)
    }
  }
  mean_diff$M_diff <- with(mean_diff, M_high - M_low)
  mean_diff
}
my_diff <- compute_diff()
head(my_diff)
tail(my_diff)
```


Plot

```{r, fig.height = 4}
ggplot(my_diff,
       aes(x = Threshold_high, y = M_diff, colour = factor(Threshold_low))) +
  geom_point() +
  geom_line(aes(group = Threshold_low))
```

We would like to pick a point that has high `M_diff`, but we also need to keep N
reasonably high. There is a fairly steep increase in the mean proficiency 
difference...

1) For a `high_profic` threshold of 27 vs 26
2) For a `low_profic` threshold of 22 vs 23

Unfortunately, the figure below shows that these cutoff points also mean
a sharp drop in the Ns:

```{r, echo = FALSE}
my_reasonable <- with(my_n, 
                      Threshold_type == "low_profic" & Threshold %in%
                        min_low_profic_thresh : (max_high_profic_thresh - 1) |
                        Threshold_type == "high_profic" & Threshold %in%
                        (min_low_profic_thresh + 1) : max_high_profic_thresh)
```


```{r, echo = FALSE, fig.height = 4}
ggplot(my_n[my_reasonable, ],
       aes(x = Threshold, y = N, colour = Condition,
                      linetype = Threshold_type, shape = Threshold_type)) +
  geom_point() +
  geom_line() +
  ylim(6, 16)
```

