---
title: "GAMM analysis"
author: "Guillermo Montero-Melis"
date: "7 februari 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include = TRUE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(mgcv)  # GAMs and GAMMs (Wood 2006)
library(itsadug)
# library(ggplot2)
# library(lme4)
# library(boot)  # for inv.logit()
# library(knitr)  # for kable()
```

```{r}
## Specify some global parameters

# infoMessages('off')  # suppresses messages from plot_smooth() / plot_diff()

# adjust figure heght/width when not going with default (espec. for 2x2 plots)
myfighe_NS_L2 <- 6
myfighe_L2_prof <- 6
myfigwi <- 7
```


```{r}
## Convenience functions

# print deviance explained as percentage
dev_expl <- function(fm) {
  devi <- summary(fm)$dev.expl
  paste0(round(100 * devi, 1), '% dev. explained')
}

# Function used to load models if they have already been saved,
# rather than fitting them anew each time the script is called
load_or_fit <- function(fm_name, fm_code, forcefit = FALSE) {
  rel_path <- './gamms/'
  # the string may contain '\n'
  fm_code <- gsub('\n', '', fm_code)
  
  # check first if there is a saved rda object with that name; if so, load it
  model_exists <- paste0(fm_name, '.rda') %in% list.files(path = rel_path)
  # if forcefit = T it will always fit the model
  if((!forcefit) & model_exists) {
    load(paste0(rel_path, fm_name, '.rda'), .GlobalEnv)
    
  } else {  # fit it and save it to right folder
    # record the time it takes to fit
    fm_time <- paste0(fm_name, ".t")
    ptm <- proc.time()
    assign(fm_name, eval(parse(text = fm_code)), envir = .GlobalEnv)
    assign(fm_time, proc.time() - ptm, envir = .GlobalEnv)
    # both the model and the time it took can be saved as same file
    save(list = c(fm_name, fm_time), file = paste0(rel_path, fm_name, '.rda'))
    # print a brief model summary when it has been fitted and time to fit
    fm <- get(fm_name)
    fm_t <- get(fm_time)
    print(fm)
    print(fm_t)
  }
}

# function to plot the differences between native and L2 speakers
plot_NS_L2 <- function(fm, primed_cond = NULL, ylim1 = c(-4, 6), ylim2 = c(-.5, 6)) {
  layout(matrix(1:4, ncol = 2, byrow=TRUE), heights = c(1.5, 1))
  par(mai = c(.7, .8, .5, 0.2))
  mycex <- 0.9
  NS_primed <- paste0('NS.', primed_cond)
  L2_primed <- paste0('L2.', primed_cond)
  # plot NS control
  plot_smooth(fm, view = 'Trial', cond = list(GroupCondit = 'NS.Control'),
              col = 'blue', lty = 2, rug = FALSE, ylim = ylim1, rm.ranef=TRUE,
              main = 'Native speakers', ylab = paste0('Log-likelihood\nof ', primed_cond, " verb"),
              hide.label = TRUE)
  # plot NS primed
  plot_smooth(fm, view = 'Trial', cond = list(GroupCondit = NS_primed),
              col = 'red', rug = FALSE, rm.ranef = TRUE, add = TRUE,
              hide.label = TRUE)
  # add legend
  legend(x = 1, y = ylim1[2], legend = c(paste0(primed_cond, '-primed'), 'Control'),
         col = c('red', 'blue'), lty = 1:2, cex = mycex, box.lty = 0)
  
  # L2 speakers Control
  plot_smooth(fm, view = 'Trial', cond = list(GroupCondit = 'L2.Control'),
              col = 'blue', lty = 2, rug = FALSE, ylim = ylim1, rm.ranef = TRUE,
              main = 'L2 speakers', ylab = paste0('Log-likelihood\nof ', primed_cond, " verb"),
              hide.label = TRUE)
  # L2 speakers primed
  plot_smooth(fm, view = 'Trial', cond = list(GroupCondit = L2_primed),
              col = 'red', rug = FALSE, rm.ranef = TRUE, add = TRUE,
              hide.label = TRUE)
  # add legend
  legend(x = 1, y = ylim1[2], legend = c(paste0(primed_cond, '-primed'), 'Control'),
         col = c('red', 'blue'), lty = 1:2, cex = mycex, box.lty = 0)
  
  # Now plot the estimated differences with itsadug::plot_diff()
  plot_diff(fm, view = 'Trial', comp = list(GroupCondit = c(NS_primed, 'NS.Control')),
            ylim = ylim2, rm.ranef=TRUE, 
            main = paste0('NS: ', primed_cond, '-primed',  ' - Control'),
            ylab = paste('Diff. in log-likelihood\nof ', primed_cond, 'verb'), hide.label = TRUE)
  plot_diff(fm, view = 'Trial', comp = list(GroupCondit = c(L2_primed, 'L2.Control')),
            ylim = ylim2, rm.ranef=TRUE, 
            main = paste0('L2: ', primed_cond, '-primed',  ' - Control'),
            ylab = paste('Diff. in log-likelihood\nof ', primed_cond, 'verb'), hide.label = TRUE)
}


# function to plot the effects by L2 speakers' proficiency
plot_L2_profic <- function(fm, primed_cond = NULL, ylim1 = c(-6, 8),
                           cloze_range = c(6, 36)) {
  cloze_scores <- seq(cloze_range[1], cloze_range[2], length.out = 6)
  # we want to have one common legend for all plots
  layout(rbind(1, matrix(2:7, ncol=3, byrow=TRUE)), heights = c(1, 6, 6))
  # add common legend
  par(mai=c(0,0,0,0))
  plot.new()
  legend(x = "center", ncol = 2, legend = c(paste0(primed_cond, '-primed'), 'Control'),
         col = c('red', 'blue'), lty = 1:2, box.lty = 0, cex = 1.5)
  # Now make plot for the model estimates at the different cloze scores in cloze_scores
  # par(mai=rep(0.2, 4))
  par(mai = c(.7, .8, .5, 0))
  for(cloze in cloze_scores) {
    plot_smooth(fm, view = 'Trial', cond = list(Condition = 'Control', ClozeScore = cloze),
                col = 'blue', lty = 2, rug = FALSE, ylim = ylim1, rm.ranef = TRUE,
                hide.label = TRUE, ylab = paste0('Log-likelihood\nof ', primed_cond, " verb"),
                main = paste('Cloze score =', cloze))
    plot_smooth(fm, view = 'Trial', cond = list(Condition = primed_cond, ClozeScore = cloze),
                col = 'red', rug = FALSE, rm.ranef = TRUE, hide.label = TRUE, add = TRUE)
    }
}
```


## Load data and process

```{r}
#  ------------------------------------------------------------------------
#  Load data
#  ------------------------------------------------------------------------

# The data is created in the script 'processing/compute_dependent_measures.R'
# There is the normal and the liberally coded version, see script for difference

getwd()

# load
d <- read.csv('../data/data_DVs.csv', fileEncoding = 'UTF-8', stringsAsFactors = TRUE)
# simplify somewhat
d <- d %>%
  select(Subject:VideoName, P_V, M_V) %>%
  rename(Trial = VideoTrial)
head(d)
str(d)

# participant data
ppts <- read.csv("../data/participants.csv", fileEncoding = "UTF-8", stringsAsFactors = TRUE)
head(ppts)
# add speakers' clozescore to d:
d <- left_join(d, ppts %>% select(Subject, ClozeScore))
head(d)
str(d)

# Subject needs to be a factor
d$Subject <- factor(d$Subject)

# the way to analyze crossed factors seems to be to combine their levels into 
# a single factor, see Wood's comment here
# http://grokbase.com/t/r/r-help/113qaadxt4/r-how-to-add-in-interaction-terms-in-gamm
d$GroupCondit <- with(d, interaction(Group, Condition))

## Data for analysis of Path-priming vs control
dp <- d %>% filter(Condition != 'Manner')
dp$GroupCondit <- factor(dp$GroupCondit, levels = c("NS.Control", "NS.Path",
                                                    "L2.Control", "L2.Path"))
# note that NS control is the reference group
contrasts(dp$GroupCondit)
## Data for analysis of Manner-priming vs control
dm <- d %>% filter(Condition != 'Path')
dm$GroupCondit <- factor(dm$GroupCondit, levels = c("NS.Control", "NS.Manner",
                                                    "L2.Control", "L2.Manner"))
# note that NS control is the reference group
contrasts(dm$GroupCondit)
```




## Do the primed conditions differ from the Control conditions?

This question is addressed separately for Path-verb priming and Manner-verb priming.

### Path-priming

Are speakers in the Path-primed condition more likely to use Path-verbs relative to the control condition?

Two crossed factors:

- Condition: Path-primed vs control
- Group: L2 speakers (L2) vs native speakers (NS)


```{r}
# # model to fit; the simpler one has only random intercepts and slopes for speakers
# # saved as string to be passed to the load_or_fit() function
# fm_p1.expr <- "bam(P_V ~ GroupCondit + s(Trial, by = GroupCondit) +
#                     s(Subject, bs = 're') + s(Trial, Subject, bs = 're') +
#                     s(VideoName, bs = 're') + s(Trial, VideoName, bs = 're'),
#                   data = dp,
#                   family = 'binomial')"
# # load or fit model
# load_or_fit("fm_p1", fm_p1.expr)
# summary(fm_p1)
```


```{r}
# ## plot
# plot_smooth(fm_p1, view = 'Trial', plot_all = 'GroupCondit', rm.ranef=TRUE)
```



```{r, fig.height = myfighe_NS_L2, fig.width = myfigwi, results = 'hide'}
# # set same y-axis for both
# myylim1 <- c(-1.5, 6)
# myylim2 <- c(-.5, 6)
# par(mfrow = c(2, 2))
# # Native speakers
# plot_smooth(fm_p1, view = 'Trial', cond = list(GroupCondit = 'NS.Control'),
#             col = 'red', rug = FALSE, ylim = myylim1, rm.ranef=TRUE,
#             main = 'NS: Path-primed vs Control')
# plot_smooth(fm_p1, view = 'Trial', cond = list(GroupCondit = 'NS.Path'),
#             col = 'blue', add = T, rm.ranef=TRUE)
# # L2 speakers
# plot_smooth(fm_p1, view = 'Trial', cond = list(GroupCondit = 'L2.Control'),
#             col = 'red', rug = FALSE, ylim = myylim1, rm.ranef=TRUE,
#             main = 'L2: Path-primed vs Control')
# plot_smooth(fm_p1, view = 'Trial', cond = list(GroupCondit = 'L2.Path'),
#             col = 'blue', add = T, rm.ranef=TRUE)
# 
# # and more to the point
# plot_diff(fm_p1, view = 'Trial', comp = list(GroupCondit = c('NS.Path', 'NS.Control')),
#           ylim = myylim2, rm.ranef=TRUE)
# plot_diff(fm_p1, view = 'Trial', comp = list(GroupCondit = c('L2.Path', 'L2.Control')),
#           ylim = myylim2, rm.ranef=TRUE)
```


```{r}
# # more complex model fits factor smooths for speakers
# fm_p2.expr <- "bam(P_V ~ GroupCondit + s(Trial, by = GroupCondit) +
#                     s(Trial, Subject, bs = 'fs'),
#                   data = dp,
#                   family = 'binomial')"
# # load model or fit
# load_or_fit("fm_p2", fm_p2.expr)
# summary(fm_p2)
```

```{r}
# par(mfrow = c(1, 1))
# ## plot
# plot_smooth(fm_p2, view = 'Trial', plot_all = 'GroupCondit', rm.ranef=TRUE)
# # show by-speaker random smooths
# plot(fm_p2, all.terms = TRUE)
```


```{r, fig.height = myfighe_NS_L2, fig.width = myfigwi, echo = FALSE, message=FALSE}
# # set same y-axes for both
# myylim1 <- c(-1.5, 5)
# myylim2 <- c(-1, 4.5)
# par(mfrow = c(2, 2))
# 
# # Native speakers
# plot_smooth(fm_p2, view = 'Trial', cond = list(GroupCondit = 'NS.Control'),
#             col = 'red', rug = FALSE, ylim = myylim1, rm.ranef=TRUE,
#             main = 'NS: Path-primed vs Control')
# plot_smooth(fm_p2, view = 'Trial', cond = list(GroupCondit = 'NS.Path'),
#             col = 'blue', rm.ranef=TRUE, add = TRUE)
# # L2 speakers
# plot_smooth(fm_p2, view = 'Trial', cond = list(GroupCondit = 'L2.Control'),
#             col = 'red', rug = FALSE, ylim = myylim1, rm.ranef=TRUE,
#             main = 'L2: Path-primed vs Control')
# plot_smooth(fm_p2, view = 'Trial', cond = list(GroupCondit = 'L2.Path'),
#             col = 'blue', rm.ranef=TRUE, add = TRUE)
# 
# # and more to the point
# plot_diff(fm_p2, view = 'Trial', comp = list(GroupCondit = c('NS.Path', 'NS.Control')),
#           ylim = myylim2, rm.ranef=TRUE)
# plot_diff(fm_p2, view = 'Trial', comp = list(GroupCondit = c('L2.Path', 'L2.Control')),
#           ylim = myylim2, rm.ranef=TRUE)
# par(mfrow = c(1, 1))
```

Previous models (not shown here) showed that random **intercepts** by items made sense, but not random slopes.
We add the former but not the latter.

Thus, the model is specified as:

`M_V ~ GroupCondit + s(Trial, by = GroupCondit) + s(Trial, Subject, bs = 'fs') + s(VideoName, bs = 're')`

Predictors are:

- `Group-Condition` interaction as a main effect
- A smooth function of `Trial` using thin plate regression splines (default)
- We allow this function to vary for each of the levels of `Group-Condition`
- Factor smooths for `Subject` to capture non-linear random effects of speakers
- Random intercepts by items (i.e., `VideoName`)


```{r}
# add random intercepts by items to fm_p2
fm_p3.expr <- "bam(P_V ~ GroupCondit + s(Trial, by = GroupCondit) +
                    s(Trial, Subject, bs = 'fs') + s(VideoName, bs = 're'),
                  data = dp,
                  family = 'binomial')"
# load model or fit
load_or_fit("fm_p3", fm_p3.expr)
summary(fm_p3)
```



```{r, results = 'hide'}
par(mfrow = c(1, 1))
## plot
plot_smooth(fm_p3, view = 'Trial', plot_all = 'GroupCondit', rm.ranef=TRUE)
# show by-speaker random smooths and by-item random intercepts
par(mfrow = c(1, 2))
plot(fm_p3, select = 5)
plot(fm_p3, select = 6)
```


```{r, fig.height = myfighe_NS_L2, fig.width = myfigwi, echo = FALSE, results = 'hide'}
plot_NS_L2(fm_p3, primed_cond = 'Path')
```


### Manner priming

```{r}
# # model to fit; the simpler one has only random intercepts and slopes for speakers
# # saved as string to be passed to the load_or_fit() function
# fm_m1.expr <- "bam(M_V ~ GroupCondit + s(Trial, by = GroupCondit) +
#                     s(Subject, bs = 're') + s(Trial, Subject, bs = 're') +
#                     s(VideoName, bs = 're') + s(Trial, VideoName, bs = 're'),
#                   data = dm,
#                   family = 'binomial')"
# # load or fit model
# load_or_fit("fm_m1", fm_m1.expr)
# summary(fm_m1)
```


```{r}
# ## plot
# plot_smooth(fm_m1, view = 'Trial', plot_all = 'GroupCondit', rm.ranef=TRUE)
```


```{r, fig.height = myfighe_NS_L2, fig.width = myfigwi}
# # set same y-axis for both
# myylim1 <- c(-6, 3.5)
# myylim2 <- c(-0.5, 8)
# par(mfrow = c(2, 2))
# # Native speakers
# plot_smooth(fm_m1, view = 'Trial', cond = list(GroupCondit = 'NS.Control'),
#             col = 'red', rug = FALSE, ylim = myylim1, rm.ranef=TRUE,
#             main = 'NS: Manner-primed vs Control')
# plot_smooth(fm_m1, view = 'Trial', cond = list(GroupCondit = 'NS.Manner'),
#             col = 'blue', rm.ranef=TRUE, add = T)
# # L2 speakers
# plot_smooth(fm_m1, view = 'Trial', cond = list(GroupCondit = 'L2.Control'),
#             col = 'red', rug = FALSE, ylim = myylim1, rm.ranef=TRUE,
#             main = 'L2: Manner-primed vs Control')
# plot_smooth(fm_m1, view = 'Trial', cond = list(GroupCondit = 'L2.Manner'),
#             col = 'blue', rm.ranef=TRUE, add = T)
# 
# # and more to the point
# plot_diff(fm_m1, view = 'Trial', comp = list(GroupCondit = c('NS.Manner', 'NS.Control')),
#           ylim = myylim2, rm.ranef=TRUE)
# plot_diff(fm_m1, view = 'Trial', comp = list(GroupCondit = c('L2.Manner', 'L2.Control')),
#           ylim = myylim2, rm.ranef=TRUE)
```



```{r}
# # more complex model fits factor smooths for speakers
# fm_m2.expr <- "bam(M_V ~ GroupCondit + s(Trial, by = GroupCondit) +
#                     s(Trial, Subject, bs = 'fs'),
#                   data = dm,
#                   family = 'binomial')"
# # load model or fit
# load_or_fit("fm_m2", fm_m2.expr)
# summary(fm_m2)
```


```{r}
# par(mfrow = c(1, 1))
# ## plot
# plot_smooth(fm_m2, view = 'Trial', plot_all = 'GroupCondit', rm.ranef=TRUE)
# # show by-speaker random smooths
# plot(fm_m2, all.terms = TRUE)
```


```{r, fig.height = myfighe_NS_L2, fig.width = myfigwi, echo = FALSE, message=FALSE}
# # set same y-axes for both
# myylim1 <- c(-4.5, 2.5)
# myylim2 <- c(-.5, 6)
# par(mfrow = c(2, 2))
# 
# # Native speakers
# plot_smooth(fm_m2, view = 'Trial', cond = list(GroupCondit = 'NS.Control'),
#             col = 'red', rug = FALSE, ylim = myylim1, rm.ranef=TRUE,
#             main = 'NS: Manner-primed vs Control')
# plot_smooth(fm_m2, view = 'Trial', cond = list(GroupCondit = 'NS.Manner'),
#             col = 'blue', rm.ranef=TRUE, add = T)
# # L2 speakers
# plot_smooth(fm_m2, view = 'Trial', cond = list(GroupCondit = 'L2.Control'),
#             col = 'red', rug = FALSE, ylim = myylim1, rm.ranef=TRUE,
#             main = 'L2: Manner-primed vs Control')
# plot_smooth(fm_m2, view = 'Trial', cond = list(GroupCondit = 'L2.Manner'),
#             col = 'blue', rm.ranef=TRUE, add = T)
# 
# # and more to the point
# plot_diff(fm_m2, view = 'Trial', comp = list(GroupCondit = c('NS.Manner', 'NS.Control')),
#           ylim = myylim2, rm.ranef=TRUE)
# plot_diff(fm_m2, view = 'Trial', comp = list(GroupCondit = c('L2.Manner', 'L2.Control')),
#           ylim = myylim2, rm.ranef=TRUE)
```


Previous models (not shown here) showed that random **intercepts** by items made sense, but not random slopes.
We add the former but not the latter.

Thus, the model is specified as:

`M_V ~ GroupCondit + s(Trial, by = GroupCondit) + s(Trial, Subject, bs = 'fs') + s(VideoName, bs = 're')`

Predictors are:

- `Group-Condition` interaction as a main effect
- A smooth function of `Trial` using thin plate regression splines (default)
- We allow this function to vary for each of the levels of `Group-Condition`
- Factor smooths for `Subject` to capture non-linear random effects of speakers
- Random intercepts by items (i.e., `VideoName`)


```{r}
# add random intercepts by items to fm_p2
fm_m3.expr <- "bam(M_V ~ GroupCondit + s(Trial, by = GroupCondit) +
                    s(Trial, Subject, bs = 'fs') + s(VideoName, bs = 're'),
                  data = dm,
                  family = 'binomial')"
# load model or fit
load_or_fit("fm_m3", fm_m3.expr)
summary(fm_m3)
```


```{r, results = 'hide'}
par(mfrow = c(1, 1))
## plot
plot_smooth(fm_m3, view = 'Trial', plot_all = 'GroupCondit', rm.ranef = TRUE)
# show by-speaker random smooths and by-item random intercepts
par(mfrow = c(1, 2))
plot(fm_m3, select = 5)
plot(fm_m3, select = 6)
```



```{r, fig.height = myfighe_NS_L2, fig.width = myfigwi, echo = FALSE, results = 'hide'}
plot_NS_L2(fm_m3, primed_cond = 'Manner')
```



## Effect of proficiency on adaptation?

Now we consider only L2 data and ask whether the time course of adaptation differs with respect to proficiency.
Following Wieling et al (2014) we model the effect of trial and proficiency using tensor products.

```{r}
# subset L2 data only
dp_l2 <- dp %>% filter(Group == "L2")
dp_l2$Condition <- factor(dp_l2$Condition)
# contain only L2 speakers in the Path/Control conditions
str(dp_l2)
dm_l2 <- dm %>% filter(Group == "L2")
dm_l2$Condition <- factor(dm_l2$Condition)
# contain only L2 speakers in the Manner/Control conditions
str(dm_l2)
```



### Path

```{r}
# fm_p_prof1.expr <- "bam(P_V ~
#                     te(Trial, ClozeScore),
#                   data = dp_l2,
#                   family = 'binomial')"
# # load model or fit
# load_or_fit("fm_p_prof1", fm_p_prof1.expr)
# summary(fm_p_prof1)
# # plot(fm_p_prof1)
```


```{r}
# fm_p_prof2.expr <- "bam(P_V ~
#                     te(Trial, ClozeScore, by = Condition),
#                   data = dp_l2,
#                   family = 'binomial')"
# # load model or fit
# load_or_fit("fm_p_prof2", fm_p_prof2.expr)
# summary(fm_p_prof2)
# # plot(fm_p_prof2, all.terms = TRUE)
```


```{r}
# fm_p_prof3.expr <- "bam(P_V ~ Condition +
#                     te(Trial, ClozeScore, by = Condition),
#                   data = dp_l2,
#                   family = 'binomial')"
# # load model or fit
# load_or_fit("fm_p_prof3", fm_p_prof3.expr)
# summary(fm_p_prof3)
# # plot(fm_p_prof3, all.terms = TRUE)
```


```{r}
# fm_p_prof4.expr <- "bam(P_V ~ Condition +
#                     te(Trial, ClozeScore, by = Condition) +
#                     s(Subject, bs = 're'),
#                   data = dp_l2,
#                   family = 'binomial')"
# # load model or fit
# load_or_fit("fm_p_prof4", fm_p_prof4.expr)
# summary(fm_p_prof4)
# # plot(fm_p_prof4, all.terms = TRUE)
```


```{r}
# fm_p_prof5.expr <- "bam(P_V ~ Condition +
#                     te(Trial, ClozeScore, by = Condition) +
#                     s(Trial, Subject, bs = 're'),
#                   data = dp_l2,
#                   family = 'binomial')"
# # load model or fit
# load_or_fit("fm_p_prof5", fm_p_prof5.expr)
# summary(fm_p_prof5)
# # plot(fm_p_prof5, all.terms = TRUE)
```


```{r}
# fm_p_prof6.expr <- "bam(P_V ~ Condition +
#                     te(Trial, ClozeScore, by = Condition) +
#                     s(Subject, bs = 're') +
#                     s(Trial, Subject, bs = 're'),
#                   data = dp_l2,
#                   family = 'binomial')"
# # load model or fit
# load_or_fit("fm_p_prof6", fm_p_prof6.expr)
# summary(fm_p_prof6)
# # plot(fm_p_prof6, all.terms = TRUE)
```


```{r}
# fm_p_prof7.expr <- "bam(P_V ~ Condition +
#                     te(Trial, ClozeScore, by = Condition) +
#                     s(Subject, bs = 're') +
#                     s(Trial, Subject, bs = 're') +
#                     s(VideoName, bs = 're'),
#                   data = dp_l2,
#                   family = 'binomial')"
# # load model or fit
# load_or_fit("fm_p_prof7", fm_p_prof7.expr)
# summary(fm_p_prof7)
# # plot(fm_p_prof7, all.terms = TRUE)
```


```{r}
# fm_p_prof8.expr <- "bam(P_V ~ Condition +
#                     te(Trial, ClozeScore, by = Condition) +
#                     s(Subject, bs = 're') +
#                     s(Trial, Subject, bs = 're') +
#                     s(VideoName, bs = 're') +
#                     s(Trial, VideoName, bs = 're'),
#                   data = dp_l2,
#                   family = 'binomial')"
# # load model or fit
# load_or_fit("fm_p_prof8", fm_p_prof8.expr)
# summary(fm_p_prof8)
# # plot(fm_p_prof8)
```



```{r}
fm_p_prof9.expr <- "bam(P_V ~ Condition +
                    te(Trial, ClozeScore, by = Condition) +
                    s(Trial, Subject, bs = 'fs') +
                    s(VideoName, bs = 're'),
                  data = dp_l2,
                  family = 'binomial')"
# load model or fit
load_or_fit("fm_p_prof9", fm_p_prof9.expr)
summary(fm_p_prof9)
```


```{r}
# show by-speaker random smooths and by-item random intercepts
par(mfrow = c(1, 2))
plot(fm_p_prof9, select = 3)
plot(fm_p_prof9, select = 4)
```


```{r, fig.height = myfighe_L2_prof, fig.width = myfigwi, results='hide'}
plot_L2_profic(fm_p_prof9, primed_cond = 'Path')
```


```{r, results='hide'}
# difference plot in 3D
plot_diff2(fm_p_prof9, view = c('Trial', 'ClozeScore'), 
           comp = list(Condition = c('Path', 'Control')),
           rm.ranef = TRUE)
```




```{r}
fm_p_prof11.expr <- "bam(P_V ~ Condition +
                    ti(Trial, by = Condition) +
                    ti(ClozeScore, by = Condition) +
                    ti(Trial, ClozeScore, by = Condition) +
                    s(Trial, Subject, bs = 'fs') +
                    s(VideoName, bs = 're'),
                  data = dp_l2,
                  family = 'binomial')"
# load model or fit
load_or_fit("fm_p_prof11", fm_p_prof11.expr)
summary(fm_p_prof11)
```

```{r}
# show by-speaker random smooths and by-item random intercepts
par(mfrow = c(1, 2))
plot(fm_p_prof11, select = 3)
plot(fm_p_prof11, select = 4)
```


```{r, fig.height = myfighe_L2_prof, fig.width = myfigwi, results='hide'}
# plot_L2_profic(fm_p_prof11, primed_cond = 'Path')
```

```{r, results='hide'}
# # difference plot in 3D
# plot_diff2(fm_p_prof11, view = c('Trial', 'ClozeScore'), 
#            comp = list(Condition = c('Path', 'Control')), rm.ranef = TRUE)
```





### Manner


```{r}
fm_m_prof9.expr <- "bam(M_V ~ Condition +
                    te(Trial, ClozeScore, by = Condition) +
                    s(Trial, Subject, bs = 'fs') +
                    s(VideoName, bs = 're'),
                  data = dm_l2,
                  family = 'binomial')"
# load model or fit
load_or_fit("fm_m_prof9", fm_m_prof9.expr)
summary(fm_m_prof9)
```

```{r}
# show by-speaker random smooths and by-item random intercepts
par(mfrow = c(1, 2))
plot(fm_m_prof9, select = 3)
plot(fm_m_prof9, select = 4)
```


```{r, fig.height = myfighe_L2_prof, fig.width = myfigwi, results='hide'}
plot_L2_profic(fm_m_prof9, primed_cond = 'Manner')
```

```{r, results='hide'}
# difference plot in 3D
plot_diff2(fm_m_prof9, view = c('Trial', 'ClozeScore'), 
           comp = list(Condition = c('Manner', 'Control')), rm.ranef = TRUE)
```


```{r}
# fm_m_prof10.expr <- "bam(M_V ~ Condition +
#                     te(Trial, ClozeScore, by = Condition) +
#                     s(Trial, Subject, bs = 'fs') +
#                     s(VideoName, bs = 're') +
#                     s(Trial, VideoName, bs = 're'),
#                   data = dm_l2,
#                   family = 'binomial')"
# # load model or fit
# load_or_fit("fm_m_prof10", fm_m_prof10.expr)
# summary(fm_m_prof10)
# plot(fm_m_prof10, all.terms = TRUE, rug = FALSE)
```



```{r}
fm_m_prof11.expr <- "bam(M_V ~ Condition +
                    ti(Trial, by = Condition) +
                    ti(ClozeScore, by = Condition) +
                    ti(Trial, ClozeScore, by = Condition) +
                    s(Trial, Subject, bs = 'fs') +
                    s(VideoName, bs = 're'),
                  data = dm_l2,
                  family = 'binomial')"
# load model or fit
load_or_fit("fm_m_prof11", fm_m_prof11.expr)
summary(fm_m_prof11)
```

```{r}
# show by-speaker random smooths and by-item random intercepts
par(mfrow = c(1, 2))
plot(fm_m_prof11, select = 3)
plot(fm_m_prof11, select = 4)
```


```{r, fig.height = myfighe_L2_prof, fig.width = myfigwi, results='hide'}
plot_L2_profic(fm_m_prof11, primed_cond = 'Manner')
```

```{r, results='hide'}
# difference plot in 3D
plot_diff2(fm_m_prof11, view = c('Trial', 'ClozeScore'), 
           comp = list(Condition = c('Manner', 'Control')), rm.ranef = TRUE)
```
