---
title: "reply_vanRij"
author: "Guillermo Montero-Melis"
date: "13 september 2017"
output:
  html_document:
    depth: 3
    number_sections: yes
    theme: default
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document:
    toc: yes

---

In this report we carry out the GAMM analyses suggested by Jacolien van Rij (emails from 12 Sept 2017).

## Libraries

```{r, include = TRUE, echo = TRUE, warning = FALSE, message = FALSE}
library(dplyr)
packageVersion('dplyr')
library(tidyr)
packageVersion('tidyr')  # useful for gather() to convert from wide to long
library(mgcv)  # GAMs and GAMMs (Wood 2006)
packageVersion('mgcv')  # GAMs and GAMMs (Wood 2006)
library(itsadug)
packageVersion('itsadug')
# library(lme4)
# packageVersion('lme4')
# library(ggplot2)
# packageVersion('ggplot2')
# library(boot)  # for inv.logit()
# packageVersion('boot')  # for inv.logit()
# library(knitr)  # for kable()
# packageVersion('knitr')  # for kable()
# library(lazyeval)  # lazy evaluation used in bysubj() function [summarise_]
# packageVersion('lazyeval')
# library(effects)
# library(xtable)
```




## Data loading and processing

```{r}
##  Load data
# The data is created in the script 'processing/compute_dependent_measures.R'
# There is the normal and the liberally coded version (see script for difference).
# Here I use the normal coding.

# load
d <- read.csv('../data/data_DVs.csv', fileEncoding = 'UTF-8', stringsAsFactors = TRUE)
# simplify somewhat
d <- d %>%
  select(Subject:VideoName, P_V, M_V) %>%
  rename(Trial = VideoTrial)
# head(d)
# str(d)

# Rename "Control" condition to "Baseline"
levels(d$Condition)[levels(d$Condition) == "Control"] <- "Baseline"

# participant data
ppts <- read.csv("../data/participants.csv", fileEncoding = "UTF-8",
                 stringsAsFactors = TRUE)
# transform ClozeScore to z-score (as.vector prevents it from becoming a matrix)
ppts$zClozeScore <- as.vector(scale(ppts$ClozeScore))
# center ClozeScore but not scaling (as.vector prevents it from becoming a matrix)
ppts$cClozeScore <- as.vector(scale(ppts$ClozeScore, scale = FALSE))

# add speakers' clozescore to d:
d <- left_join(d, ppts %>% select(Subject, ClozeScore, zClozeScore, cClozeScore))

# Subject needs to be a factor
d$Subject <- factor(d$Subject)

# in Group, let the native speakers (NS) be the reference group
d$Group <- factor(d$Group, levels = c('NS', 'L2'))

head(d)
str(d)
```


## Source some convenience functions

```{r}
## Source somewhat more complex functions

# source multiplot function
source("functions/multiplot_fnc.R")

# Function used to load models if they have already been saved,
# rather than fitting them anew each time the script is called
source("functions/load_or_fit_fnc.R")

# Two functions to a) plot the differences between NS and L2 speakers from GAMMs,
# and b) plot the effects by L2 speakers' proficiency from GAMMs
source("functions/plot_gams_fnc.R")
```



# Guillermo's original analyses

## Data for GAMs

- We will subset the data so that we can run one model for native speakers and another one for L2 speakers
- In each model we compare Baseline Path vs Primed Path and Baseline Manner vs Exposed Manner

Below: Subset data, define factors, set coding scheme, etc.

```{r}
## Single factor to model VerbType * Condition, using dummy coding:
# (The way to analyze crossed factors is to combine their levels into 
# a single factor, see Wood's comment here
# http://grokbase.com/t/r/r-help/113qaadxt4/r-how-to-add-in-interaction-terms-in-gamm)

# First need to convert data to long format:
d_long <- gather(d, VerbType, Used, P_V:M_V)
# single factor
d_long$VbType_Cond <- with(d_long, interaction(VerbType, Condition))

# For the subjects  to be properly fitted as random smooths in the GAMs
# we have to "pretend" that a baseline subject is two different subjects,
# one for the comparison to path-primed, the other to manner-primed participants;
# this may not be ideal statistically, but not doing this would turn the random
# smooths into random intercepts, which is much worse.
d_long$Subject <- with(d_long, interaction(Subject, VerbType))

## Subset data:
# Data for analysis of Native speakers (note we are removing observations that
# correspond to path verbs produced in the manner-primed condition or to manner
# verbs produced in the path-primed condition)
d_ns <- d_long %>% filter(Group == "NS" & ! VbType_Cond %in% c("P_V.Manner", "M_V.Path"))
# drop unused factors for subject (I think this is important for gam fitting)
d_ns$Subject <- factor(d_ns$Subject)
# Data for analysis of L2 speakers
d_l2 <- d_long %>% filter(Group == "L2" & ! VbType_Cond %in% c("P_V.Manner", "M_V.Path"))
d_l2$Subject <- factor(d_l2$Subject)

# Order levels correctly in each data frame
d_ns$VbType_Cond <- factor(d_ns$VbType_Cond, levels = c("P_V.Baseline", "P_V.Path",
                                                      "M_V.Baseline", "M_V.Manner"))
d_l2$VbType_Cond <- factor(d_l2$VbType_Cond, levels = c("P_V.Baseline", "P_V.Path",
                                                      "M_V.Baseline", "M_V.Manner"))
# note that log-odds of path verbs in baseline condition becomes the reference
# level in both groups (NS and L2 speakers)
contrasts(d_ns$VbType_Cond)
contrasts(d_l2$VbType_Cond)
```


## Differences in adaptation over time (trial-by-trial analysis using GAMs)


### Native speakers -- trial-by-trial

Previous models (not shown here) showed that random *intercepts* by items made sense, but not random slopes.
We add the former but not the latter.

Thus, the model is specified as:

`PathVerb ~ VbType_Cond + s(Trial, by = VbType_Cond) + s(Trial, Subject, bs = 'fs') + s(VideoName, bs = 're')`

Predictors are:

- `VbType_Cond`: Group-Condition interaction as a fixed effect
- `s(Trial, by = VbType_Cond)`: A smooth function of `Trial` allowing the function to differ for each level of Group-Condition (thin plate regression splines, the default)
- `s(Trial, Subject, bs = 'fs')`: Factor smooths for `Subject` to capture non-linear random effects of speakers
- `s(VideoName, bs = 're')`: Random intercepts by items (i.e., `VideoName`)


```{r gam_ns}
# the expression that is passed to load_or_fit()
gam_ns.expr <- "bam(Used ~ VbType_Cond + s(Trial, by = VbType_Cond) +
                    s(Trial, Subject, bs = 'fs') + s(VideoName, bs = 're'),
                  data = d_ns,
                  family = 'binomial')"
# load model or fit
load_or_fit("gam_ns", gam_ns.expr)
```

Summary of the model:

```{r}
summary(gam_ns)
```



Significance of the different terms in the model:

```{r}
anova(gam_ns)
```


Plot model estimates by condition, random smooth adjustments by speakers and QQ-plot of random by-item intercepts:

```{r, results = 'hide'}
## plot
plot_smooth(gam_ns, view = 'Trial', plot_all = 'VbType_Cond', rm.ranef=TRUE)
```


```{r}
# show by-speaker random smooths and by-item random intercepts
par(mfrow = c(1, 2))
plot(gam_ns, select = 5)
plot(gam_ns, select = 6)
```



Plot differences between conditions (model estimates):

```{r}
plot_gam_main(gam_ns, "NS")  # choose this on Linux machine
# on my old windows computer:
# plot_gam_main(gam_ns, "NS", mark.diff = FALSE)  # use of mark.diff is a hack to
# avoid an error that will stop the whole thing when I compile it on my work pc
```





# Following J van Rij's suggestions

I directly go to JvR's 3rd suggestion; quote:

```
[...] So a better idea is probably to go completely binary:

dat$IsPrime <- ifelse(dat$Condition == "Prime", 1, 0)
dat$IsPath   <- ifelse(dat$Group == "Path-verb", 1, 0)
dat$IsPathAndPrime <- ifelse(dat$Condition == "Prime" & dat$Group ==  "Path-verb", 1, 0)

# 3-way interaction model:
DV ~ s(Trial) + s(Trial, by=IsPrime) + s(Trial, by=IsPath) + s(Trial, by=IsPathAndPrime)
# only 2-way interactions:
DV ~ s(Trial) + s(Trial, by=IsPrime) + s(Trial, by=IsPath)

The partial effect of the interaction term ("the difference of the difference") could be visualized by mgcv's plot() function, or by itsadug's get_modelterm() and plot_error() for more flexibility (see examples in help file of get_modelterm()).
```

## Data coding


```{r}
# Original coding of data involves two separate binary DVs for each description:
# P_V = was a path verb used (yes/no); M_V = was a manner verb used (yes/no)
head(d, 3)

# Instead of two separate DVs for the use of Path and Manner verbs, create an
# indicator variable VerbType and let the DV (=Used) denote whether that verb
# type was used (=1) or not (=0)
d_jvr <- gather(d, VerbType, Used, P_V:M_V)

# Drop unused variables to simplify somewhat
d_jvr <- d_jvr %>% select(-(ClozeScore:cClozeScore))

# Apply JvR's coding scheme:
d_jvr$IsPrimed <- ifelse(d_jvr$Condition == "Baseline", 0, 1)
d_jvr$IsManner <- ifelse(d_jvr$VerbType == "P_V", 0, 1)
d_jvr$IsMannerAndPrimed <- with(d_jvr, ifelse(IsPrimed == 1 & IsManner ==  1, 1, 0))

# We remove observations that correspond to path verbs produced in the manner-
# primed condition or to manner verbs produced in the path-primed condition
d_jvr <- d_jvr %>% filter(! ((Condition == "Path" & VerbType == "M_V") | 
                           Condition == "Manner" & VerbType == "P_V"))

# this is how we have defined our conditions
unique(d_jvr[, c("Condition", "VerbType", "IsPrimed", "IsManner", "IsMannerAndPrimed")])

# For the subjects in baseline condition to be properly fitted as random smooths
# in the GAMswe have to "pretend" that a baseline subject is two different
# subjects, one for the comparison to path-primed, the other to manner-primed
# participants; this may not be ideal statistically, but not doing this would
# turn the random smooths into random intercepts, which is worse.
d_jvr$Subject <- with(d_jvr, interaction(Subject, VerbType))

# Subset data for native speakers only and for L2 learners only
d_jvr_ns <- d_jvr %>% filter(Group == "NS")
d_jvr_ns$Subject <- factor(d_jvr_ns$Subject)  # drop unused levels
d_jvr_l2 <- d_jvr %>% filter(Group == "L2")
d_jvr_l2$Subject <- factor(d_jvr_l2$Subject)  # drop unused levels

head(d_jvr_ns)
str(d_jvr_ns)
summary(d_jvr_ns)
```


## GAMMs Native speakers

### 3-way interaction model:

```{r}
# the expression that is passed to load_or_fit()
gam_ns_jvr_3way.expr <- "bam(Used ~
s(Trial) + s(Trial, by = IsPrimed) + s(Trial, by = IsManner) + s(Trial, by = IsMannerAndPrimed) +
s(Trial, Subject, bs = 'fs', m = 1) + s(Trial, VideoName, bs = 'fs', m = 1),
data = d_jvr_ns,
family = 'binomial')"
# load model or fit
load_or_fit("gam_ns_jvr_3way", gam_ns_jvr_3way.expr)
```


The model issues the following warning when fitted:

```
## model has repeated 1-d smooths of same variable.
```
The model summary suggests that the critical 3-way interaction is not significant, see coefficient `s(Trial):IsMannerAndPrimed`
(this will be confirmed when plotting its estimate below):

```{r}
summary(gam_ns_jvr_3way)
```

The fourth plot shows the interaction term `s(Trial):IsMannerAndPrimed`. Its 95% CI overlaps with zero, i.e. there is no adjustment to be made to the MannerAndPrimed condition beyond the additive effect of IsPrimed and IsManner.

```{r}
plot.gam(gam_ns_jvr_3way)
```

Now, when trying to visualize the estimated effects using `get_modelterm()` and `plot_error()`, the result suggests something is wrong with this model. I don't know how to make sense of the following plot:

```{r}
# # Get list with predictions:
# p <- get_modelterm(gam_ns_jvr_3way, select = 4)
# emptyPlot(range(p$terms), range(p$fit), h=0)
# plot_error(p$terms, p$fit, p$se.fit, shade=TRUE, xpd=TRUE)

# Get list with predictions and plot each of the coefficients
# Repeat the same thing with coefficients 1 through 3
for(i in 1:4) {
  p <- get_modelterm(gam_ns_jvr_3way, select = i)
  emptyPlot(range(p$terms), range(p$fit), h=0)
  plot_error(p$terms, p$fit, p$se.fit, shade=TRUE, xpd=TRUE)
}


```

Plot the random effects (following the example in `?get_modelterm`):

```{r}
# Plot random effects in separate panels:
pp <- get_modelterm(gam_ns_jvr_3way, select=5, as.data.frame=TRUE)

require(lattice)
lattice::xyplot(fit~Trial|Subject, 
                data=pp, type="l",
                xlab="Trial", ylab="Partial effect")
```

We can see that counting Subjects in the baseline condition twice, once for P_V and once for M_V, might in fact lead to redundancy in the data. Here we plot a selection and it seems like the estimated random effects for a subject for P_V are the mirror image of their estimated random effects for M_V:

```{r}
# Plot selection of subjects:
pp <- get_modelterm(gam_ns_jvr_3way, select=5,
                    cond=list(Subject=c(
                      'NS_186.P_V', 'NS_186.M_V', 'NS_187.P_V', 'NS_187.M_V',
                      'NS_188.P_V', 'NS_188.M_V', 'NS_190.P_V', 'NS_190.M_V',
                      'NS_160.P_V', 'NS_160.M_V')),
                    as.data.frame=TRUE)
lattice::xyplot(fit~Trial|Subject, 
                data=pp, type="l",
                xlab="Trial", ylab="Partial effect")
```



The autocorrelation of the residuals in the model seems not to be great:

```{r}
acf(resid(gam_ns_jvr_3way))
```


### 2-way interaction model:


```{r}
# the expression that is passed to load_or_fit()
gam_ns_jvr_2way.expr <- "bam(Used ~
s(Trial) + s(Trial, by = IsPrimed) + s(Trial, by = IsManner) +
s(Trial, Subject, bs = 'fs', m = 1) + s(Trial, VideoName, bs = 'fs', m = 1),
data = d_jvr_ns,
family = 'binomial')"
# load model or fit
load_or_fit("gam_ns_jvr_2way", gam_ns_jvr_2way.expr)
```


The model issues the following warning when fitted:

```
## model has repeated 1-d smooths of same variable.
```

```{r}
summary(gam_ns_jvr_2way)
```

Plot estimates:

```{r}
plot.gam(gam_ns_jvr_2way)
```

Again, when trying to visualize the estimated effects using `get_modelterm()` and `plot_error()`, the result suggests something is wrong with the model:

```{r}
# Repeat the same thing with coefficients 1 through 3
for(i in 1:3) {
  p <- get_modelterm(gam_ns_jvr_2way, select = i)
  emptyPlot(range(p$terms), range(p$fit), h=0)
  plot_error(p$terms, p$fit, p$se.fit, shade=TRUE, xpd=TRUE)
}
```

Plot the random effects (following the example in `?get_modelterm`):

```{r}
# Plot random effects in separate panels:
pp <- get_modelterm(gam_ns_jvr_2way, select=4, as.data.frame=TRUE)

require(lattice)
lattice::xyplot(fit~Trial|Subject, 
                data=pp, type="l",
                xlab="Trial", ylab="Partial effect")
```


The autocorrelation of the residuals in the model seems not to be great:

```{r}
acf(resid(gam_ns_jvr_2way))
```


### Model comparison

Does the model with the 3-way interaction provide a significantly better fit than the one without it? No!

```{r}
itsadug::compareML(gam_ns_jvr_3way, gam_ns_jvr_2way) 
```



## GAMMs L2 learners

### 3-way interaction model:

```{r}
# the expression that is passed to load_or_fit()
gam_l2_jvr_3way.expr <- "bam(Used ~
s(Trial) + s(Trial, by = IsPrimed) + s(Trial, by = IsManner) + s(Trial, by = IsMannerAndPrimed) +
s(Trial, Subject, bs = 'fs', m = 1) + s(Trial, VideoName, bs = 'fs', m = 1),
data = d_jvr_l2,
family = 'binomial')"
# load model or fit
load_or_fit("gam_l2_jvr_3way", gam_l2_jvr_3way.expr)
```


The model issues the following warning when fitted:

```
## Warning in gam.side(sm, X, tol = .Machine$double.eps^0.5): model has
## repeated 1-d smooths of same variable.
```
The model summary suggests that the critical 3-way interaction is not significant. This is confirmed when plotting its estimate below.

```{r}
summary(gam_l2_jvr_3way)
```

The fourth plot shows the interaction term. Its 95% CI overlaps with zero, i.e. there is no adjustment to be made to the MannerAndPrimed condition beyond the additive effect of IsPrimed and IsManner.

```{r}
plot.gam(gam_l2_jvr_3way)
```

Now, when trying to visualize the estimated effects using `get_modelterm()` and `plot_error()`, the result suggests something is wrong with this model:

```{r}
# Get list with predictions:
p <- get_modelterm(gam_l2_jvr_3way, select = 4)
emptyPlot(range(p$terms), range(p$fit), h=0)
plot_error(p$terms, p$fit, p$se.fit, shade=TRUE, xpd=TRUE)
```


The autocorrelation of the residuals in the model again seems low, although somewhat greater than for the native speaker model:

```{r}
acf(resid(gam_l2_jvr_3way))
```


### 2-way interaction model:


```{r}
# the expression that is passed to load_or_fit()
gam_l2_jvr_2way.expr <- "bam(Used ~
s(Trial) + s(Trial, by = IsPrimed) + s(Trial, by = IsManner) +
s(Trial, Subject, bs = 'fs', m = 1) + s(Trial, VideoName, bs = 'fs', m = 1),
data = d_jvr_l2,
family = 'binomial')"
# load model or fit
load_or_fit("gam_l2_jvr_2way", gam_l2_jvr_2way.expr)
```


The model issues the following warning when fitted:

```
## model has repeated 1-d smooths of same variable.
```

```{r}
summary(gam_l2_jvr_2way)
```

Plot estimates:

```{r}
plot.gam(gam_l2_jvr_2way)
```

Again, when trying to visualize the estimated effects using `get_modelterm()` and `plot_error()`, the result suggests something is wrong with the model:

```{r}
# Repeat the same thing with coefficients 1 through 3
for(i in 1:3) {
  p <- get_modelterm(gam_l2_jvr_2way, select = i)
  emptyPlot(range(p$terms), range(p$fit), h=0)
  plot_error(p$terms, p$fit, p$se.fit, shade=TRUE, xpd=TRUE)
}
```

Plot the random effects (following the example in `?get_modelterm`):

```{r}
# Plot random effects in separate panels:
pp <- get_modelterm(gam_l2_jvr_2way, select=4, as.data.frame=TRUE)

require(lattice)
lattice::xyplot(fit~Trial|Subject, 
                data=pp, type="l",
                xlab="Trial", ylab="Partial effect")
```


The autocorrelation of the residuals in the model seems not to be great:

```{r}
acf(resid(gam_l2_jvr_2way))
```



### Model comparison

Does the model with the 3-way interaction provide a significantly better fit than the one without it? No!

```{r}
itsadug::compareML(gam_l2_jvr_3way, gam_l2_jvr_2way) 
```



# Session info

```{r}
sessionInfo()
```


