---
title: "Self-priming"
author: '[Guillermo Montero-Melis](https://www.mpi.nl/people/montero-melis-guillermo)'
date: '`r as.character(format(Sys.Date(), format="%d/%m/%Y"))`'
output:
  html_document:
    depth: 2
    number_sections: yes
    theme: default
    toc: yes
---


Introduction
============

We look into possible effects of self-priming, as touched upon by reviewers
1 and 2


Reviewer concerns
-------------------

R1:

"5. Strictly speaking, each priming trial consists of two different types of
utterances: the written and orally repeated sentence, and the sentence
produced by the participant to describe the picture. I take it that in the
adaptation analysis, only the number of trials has been taken into
consideration (that is, each prime+target pair is counted as 1). Would the
result be any different if in addition to the number of primes, also the
number of path or manner constructions produced as descriptions by the
participant were taken into consideration as well? (I am thinking of a paper
by Jacobs et al. https://psyarxiv.com/brcv2/ for an analysis taking into
account self-priming)"



Our approach 
------------

Following Florian's suggestion:

Repeat the analysis while substituting Trial by:

a) The number of times the structure has been used in just production (only
   production primes) or
b) Trial + the number of times it has been used in production (both p+c
   primes).

See whether the results change and report it in the letter.

From Florian's email 2019-02-03:

"See p.77-78 in Jaeger and Snider (2013), Section 5.4. It is directly related
to this point and also contains language that might be part of our reply. On
p.78 it describes how we repeated our test (in our case, the calculation of
surprisal, in your case how you count 'trial')."



Libraries etc.
=============

```{r}
knitr::opts_chunk$set(echo = TRUE)
# Libraries and functions -------------------------------------------------

library(dplyr)
library(tidyr)  # useful for gather() to convert from wide to long
library(mgcv)     # v. 1.8-17
library(itsadug)  # v. 2.2
library(ggplot2)

source("../1901_data-script-sharing/load_or_fit_fnc.R")
# path where to look for the models and save them if fitted
model_path <- '../1901_data-script-sharing/fitted_models/'

# Two functions to a) plot adaptation effects for native and L2 speakers from
# GAMMs, and b) plot the effects by L2 speakers' proficiency.
source("../1901_data-script-sharing/plot_gams_fnc.R")
```


```{r}
fighe_l2 <- 3
figwi_l2 <- 9
```


Data
====

```{r}
# Data --------------------------------------------------------------------

# load data
d <- read.csv('../1901_data-script-sharing/data.csv', fileEncoding = 'UTF-8', 
              stringsAsFactors = TRUE) 

# Make sure Subject is a factor
d$Subject <- factor(d$Subject)
# in Group, let the native speakers (NS) be the reference group
d$Group <- factor(d$Group, levels = c('NS', 'L2'))
head(d, 3)

# Convert data to long format: each description will now have two rows, one for
# each value of VerbType (P_V = Path verb; M_V = Manner verb); note that these
# two rows were previously two different columns showing whether what was 
# encoded in the main verb in a given description encoded was path, manner,
# both or neither (see data frame d above).
d_long <- gather(d, VerbType, Used, P_V : M_V)

# Create copy of long data set, which we will further process for model fitting
d_mod <- d_long

# Combine VerbType (Path/Manner) and Condition (baseline/primed) into a single
# factor -- this is needed for the GAMMs. (The way to analyze crossed factors
# is to combine their levels into a single factor, see Wood's comment here:
# http://grokbase.com/t/r/r-help/113qaadxt4/r-how-to-add-in-interaction-terms-in-gamm)
d_mod$VbType_Cond <- with(d_mod, interaction(VerbType, Condition))	

# In order to make the relevant comparisons (see point 1 here above), baseline
# participants have to be compared twice, once for their path verb use to path-
# primed participants, once for their manner verb use to manner-primed 
# participants. This means that each of the two rows defined by VerbType will
# be used in one of the comparisons. The following renaming allows us to
# fit random by-subject smooths for each of those comparisons:
d_mod$Subject <- with(d_mod, interaction(Subject, VerbType))	

# Subset data for model fitting:
# We remove observations that	contain irrelevant comparisons given our research
# question, namely those that correspond to path verbs produced in the manner-
# primed condition or to manner	verbs produced in the path-primed condition.
d_mod <- d_mod %>% dplyr::filter(! VbType_Cond %in% c("P_V.Manner", "M_V.Path"))
# drop unused factor levels and order levels correctly
d_mod$VbType_Cond <- factor(d_mod$VbType_Cond, 
                            levels = c("P_V.Baseline", "P_V.Path", "M_V.Baseline",
                                       "M_V.Manner"))
# drop unused factors for subject (important for random effects estimation)
d_mod$Subject <- factor(d_mod$Subject)
# Condition recoded as a binary factor (Path/Manner become "Primed")
d_mod$Condition_bin <- d_mod$Condition
levels(d_mod$Condition_bin)[levels(d_mod$Condition_bin) %in% c("Path", "Manner")] <- "Primed"	
```


Add participant's cumulative productions as a predictor:

```{r}
# Add number of produced structures as predictor --------------------------

# The cumulative number of productions is equal to the number of times the target
# structure # was used up until the last trial. This is what the following function
# comptues, where v is the vector of a subject's uses of the structure ordered by trial:
my_cumsum <- function(v) {
  out <- c(0, cumsum(v)[ 1 : (length(v) - 1) ])
  out
}

# Add the following predictors:
# a) prod_primes: number of times the structure has been used in just production 
# b) prod_compr: Trial + the number of times it has been used in production. 
d_mod <- d_mod %>%
  group_by(Subject) %>%
  mutate(prod_primes = my_cumsum(Used),
         prod_compr = prod_primes + Trial)
head(data.frame(d_mod))
```


Correlations between Trial and own productions
==========================================

Visually:

```{r}
# Correlations ------------------------------------------------------------

ggplot(d_mod, aes(x = Trial, y = prod_primes)) +
  geom_jitter(alpha = .4) +
  facet_grid(VbType_Cond ~ Group) +
  geom_smooth(method = "lm")

# showing individual speakers as lines
ggplot(d_mod, aes(x = Trial, y = prod_primes)) +
  geom_jitter(alpha = .4) +
  facet_grid(VbType_Cond ~ Group) +
  geom_smooth(method = "lm") +
  geom_line(aes(group = Subject, colour = Subject)) +
  theme(legend.position = "none") 
```

Numerically (Pearson correlations)

```{r}
by(d_mod, list(d_mod$Group, d_mod$VbType_Cond), 
   function(df) cor.test(df$Trial, df$prod_primes))
```


Q1 re-analysis
==============

## Subset data

```{r}
# Subset data to native speaker data only
d_ns <- d_mod %>% dplyr::filter(Group == "NS")
# drop unused factors for subject (important for gam fitting)
d_ns$Subject <- factor(d_ns$Subject)
# note log-odds of path verbs in baseline condition becomes the reference level
contrasts(d_ns$VbType_Cond)
```


## Model with Trial only (= model reported in the paper)

```{r, results = 'hide'}
## Model with Trial only [= model reported in the paper]

# the expression that is passed to load_or_fit()
gam_ns.expr <- 
  "bam(Used ~ VbType_Cond + s(prod_primes, by = VbType_Cond) + s(prod_primes, Subject, bs = 'fs') +
s(VideoName, bs = 're'),
data = d_ns,
family = 'binomial')"
# load model or fit (fitting took about 4.5 minutes on my laptop)
load_or_fit("gam_ns", gam_ns.expr, alternPath = model_path)
# plot
plot_gam_main(gam_ns, "NS")  # choose this on Linux machine
```


## Model with self-priming from own production only


```{r, results = 'hide'}
## Model with self-priming in production only

# the expression that is passed to load_or_fit()
gam_ns_prod.expr <- 
  "bam(Used ~ VbType_Cond + s(prod_primes, by = VbType_Cond) + s(prod_primes, Subject, bs = 'fs') +
s(VideoName, bs = 're'),
data = d_ns,
family = 'binomial')"
# load model or fit (fitting took about 4.5 minutes on my laptop)
load_or_fit("gam_ns_prod", gam_ns_prod.expr, alternPath = model_path)

plot_gam_main(gam_ns_prod, "NS", show_xaxis = 'prod_primes')
```


## Model with self-priming in production and comprehension priming

```{r, results = 'hide'}
## Model with self-priming in production and comprehension priming
# the expression that is passed to load_or_fit()
gam_ns_prod_compr.expr <- 
  "bam(Used ~ VbType_Cond + s(prod_compr, by = VbType_Cond) + s(prod_compr, Subject, bs = 'fs') +
s(VideoName, bs = 're'),
data = d_ns,
family = 'binomial')"
# load model or fit (fitting took about 4.5 minutes on my laptop)
load_or_fit("gam_ns_prod_compr", gam_ns_prod_compr.expr, alternPath = model_path)

plot_gam_main(gam_ns_prod_compr, "NS", show_xaxis = 'prod_compr')
```


Q2 re-analysis
==============

## Subset data

```{r}
# Subset data to L2 speaker data only
d_l2 <- d_mod %>% dplyr::filter(Group == "L2")
# drop unused factors for subject (I think this is important for gam fitting)
d_l2$Subject <- factor(d_l2$Subject)
# note log-odds of path verbs in baseline condition becomes the reference level
contrasts(d_l2$VbType_Cond)
```


## Model with Trial only [= model reported in the paper]

```{r, results = 'hide'}
## Model with Trial only [= model reported in the paper]

# the expression that is passed to load_or_fit()
gam_l2.expr <- 
  "bam(Used ~ VbType_Cond + s(prod_primes, by = VbType_Cond) + s(prod_primes, Subject, bs = 'fs') +
s(VideoName, bs = 're'),
data = d_l2,
family = 'binomial')"
# load model or fit (fitting took about 4.5 minutes on my laptop)
load_or_fit("gam_l2", gam_l2.expr, alternPath = model_path)

plot_gam_main(gam_l2, "L2")  # choose this on Linux machine
```


## Model with self-priming in production only

```{r, results = 'hide'}
## Model with self-priming in production only

# the expression that is passed to load_or_fit()
gam_l2_prod.expr <- 
  "bam(Used ~ VbType_Cond + s(prod_primes, by = VbType_Cond) + s(prod_primes, Subject, bs = 'fs') +
s(VideoName, bs = 're'),
data = d_l2,
family = 'binomial')"
# load model or fit (fitting took about 4.5 minutes on my laptop)
load_or_fit("gam_l2_prod", gam_l2_prod.expr, alternPath = model_path)

plot_gam_main(gam_l2_prod, "L2", show_xaxis = 'prod_primes')
```


## Model with self-priming in production and comprehension priming

```{r, results = 'hide'}
## Model with self-priming in production and comprehension priming

# the expression that is passed to load_or_fit()
gam_l2_prod_compr.expr <- 
  "bam(Used ~ VbType_Cond + s(prod_compr, by = VbType_Cond) + s(prod_compr, Subject, bs = 'fs') +
s(VideoName, bs = 're'),
data = d_l2,
family = 'binomial')"
# load model or fit (fitting took about 4.5 minutes on my laptop)
load_or_fit("gam_l2_prod_compr", gam_l2_prod_compr.expr, alternPath = model_path)

plot_gam_main(gam_l2_prod_compr, "L2", show_xaxis = 'prod_compr')
```


Q3 re-analysis
==============

We use the same data as in Q2.


## Model with Trial only [= model reported in the paper]

```{r, results = 'hide', fig.height = fighe_l2, fig.width = figwi_l2}
## Model with Trial only [= model reported in the paper]

gam_l2prof.expr <- 
  "bam(Used ~ VbType_Cond + te(Trial, ClozeScore, by = VbType_Cond) +
s(Trial, Subject, bs = 'fs') + s(VideoName, bs = 're'),
data = d_l2,
family = 'binomial')"
# load model or fit (fitting took about 6.5 minutes on my laptop)
load_or_fit("gam_l2prof", gam_l2prof.expr, alternPath = model_path)

# Plot effects by condition
plot_L2_profic(gam_l2prof, primed_cond = 'Path')
plot_L2_profic(gam_l2prof, primed_cond = 'Manner')

# Plot differences between conditions (model estimates):
plot_L2_profic_diff(gam_l2prof, primed_cond = "Path")
plot_L2_profic_diff(gam_l2prof, primed_cond = "Manner")
```


## Model with self-priming in production only


```{r, results = 'hide', fig.height = fighe_l2, fig.width = figwi_l2}
## Model with self-priming in production only

# the expression that is passed to load_or_fit()
gam_l2prof_prod.expr <- 
  "bam(Used ~ VbType_Cond + te(prod_primes, ClozeScore, by = VbType_Cond) +
s(prod_primes, Subject, bs = 'fs') + s(VideoName, bs = 're'),
data = d_l2,
family = 'binomial')"
# load model or fit (fitting took about X minutes on my laptop)
load_or_fit("gam_l2prof_prod", gam_l2prof_prod.expr, alternPath = model_path)

# Plot effects by condition
plot_L2_profic(gam_l2prof_prod, primed_cond = 'Path', show_xaxis = 'prod_primes')
plot_L2_profic(gam_l2prof_prod, primed_cond = 'Manner', show_xaxis = 'prod_primes')

# Plot differences between conditions (model estimates):
plot_L2_profic_diff(gam_l2prof_prod, primed_cond = "Path", show_xaxis = 'prod_primes')
plot_L2_profic_diff(gam_l2prof_prod, primed_cond = "Manner", show_xaxis = 'prod_primes')
```


## Model with self-priming in production and comprehension priming

```{r, results = 'hide', fig.height = fighe_l2, fig.width = figwi_l2}

## Model with self-priming in production and comprehension priming

# the expression that is passed to load_or_fit()
gam_l2prof_prod_compr.expr <- 
  "bam(Used ~ VbType_Cond + te(prod_compr, ClozeScore, by = VbType_Cond) +
s(prod_compr, Subject, bs = 'fs') + s(VideoName, bs = 're'),
data = d_l2,
family = 'binomial')"
# load model or fit (fitting took about X minutes on my laptop)
load_or_fit("gam_l2prof_prod_compr", gam_l2prof_prod_compr.expr, alternPath = model_path)

# Plot effects by condition
plot_L2_profic(gam_l2prof_prod_compr, primed_cond = 'Path', show_xaxis = 'prod_compr')
plot_L2_profic(gam_l2prof_prod_compr, primed_cond = 'Manner', show_xaxis = 'prod_compr')

# Plot differences between conditions (model estimates):
plot_L2_profic_diff(gam_l2prof_prod_compr, primed_cond = "Path", show_xaxis = 'prod_compr')
plot_L2_profic_diff(gam_l2prof_prod_compr, primed_cond = "Manner", show_xaxis = 'prod_compr')
```

